WSClean: an implementation of a fast, generic wide-field imager

### 2014-01-31 articles/2014/Article-WSClean/wsclean.tex
Progress on article

\subsection{Discretization of w} \label{sec:gridding-w}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal TODO in contrast to some article? -- could be explained by having more samples at smaller $w$s. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. [TODO table of arrays with $w$ values at zenith/more angles/more freq?] For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that, in the case of the MWA, tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in section TODO.

@@

\begin{table}[h]
 \caption{Scaling of the computational cost for various imaging steps} \label{tbl:computational-cost-per-operation}
 \begin{tabular}{lll}
   \textbf{Operation} & \textbf{W-stacking} & \textbf{W-projection} \\
   \hline\hline
   Fourier transform(s) & $N_\textrm{wlay} N^2_\textrm{pix} \log N_\textrm{pix}$ & $N^2_\textrm{pix} \log N_\textrm{pix}$ \\
   W-term corrections   & $N_\textrm{wlay} N^2_\textrm{pix}$ & $N_\textrm{vis} N^2_\textrm{wkern}$ \\
   Gridding & $N_\textrm{vis}N^2_\textrm{kern}$ & $N_\textrm{vis} N_\textrm{kern}^2$ \\
   \hline\hline
 \end{tabular}
\end{table}

@@

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and minor Clean iterations (peak-finding and image subtraction). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations are so fast that performing the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, is not so relevant. However, the PSF varies for imaging with a non-zero w-value, and subtracting a constant PSF in image space leads to inaccuracies. After a number of these minor iterations, it is therefore beneficial to invert the model back to the visibilities and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

\subsection{Full-sky imaging}
For low-frequency telescopes, it can be useful to do full-sky imaging. In the case of MWA, the tile beam can have strong sidelobes at a distance of more than 90\degree from the pointing centre. Imaging these sidelobes might be relevant because it might either be scientifically interesting (e.g. when searching for transients), or be useful for self-calibration or full-sky deconvolution. For example, the latter is useful when a strong source in a sidelobe affects the sensitivity near the pointing centre.

To image the full sky, an observation is split into short snapshots, their phase centre is changed to zenith and subsequently a suitable area is imaged. Depending on desired resolution and dynamic range, this might require very large images. To make an image at the MWA resolution, the image needs to have approximately 10K pixels in width and height. The $w$-values will be very small at zenith, because only the differential heights of the antennas will contribute to $w$-value at zenith. For the MWA, the maximum height difference between tiles is $8.5$m. The MWA tiles are in fact on a slight slope, and by fitting a plane to the antennas and changing the phase centre towards the normal of that plane, the maximum $w$-value decreases to $5.5\textrm{m} / \lambda$. In the following sections, when discussing MWA zenith direction we will refer to this optimal $w$-direction.

\subsection{W-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with WSClean using the Cotton-Schwab clean algorithm. No beam correction was applied. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
To get optimal imaging performance, the $w$-values can be kept small by phase-changing towards zenith. As shown, all-sky imaging can be quite efficient with this technique. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \cdot \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. By doing the inverse corrections during visibility prediction, such a shifted visibility set can be cleaned with Cotton-Schwab iterations similar to a non-shifted set.

@@

Another method to shift an image, is by using the periodicity of the FFT function. Since sources outside the field of view will be aliased back into the field, the image can be transformed to have the alias in the centre. This complicates gridding during imaging, because the antialiasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method.

\section{Analysis} \label{sec:analysis}
We will now analyse the performance and accuracy of our implementation, and compare it with the w-projection implementation in CASA. For the analysis, we use imaging parameters which are common for MWA Imaging. When a specific parameter value is not mentioned, the settings from Table~\ref{tbl:default-parameters} are used. The number of $w$-projection planes in w-projection is kept equal to the number of $w$-layers in w-stacking, and is set to the right hand side of Eq.~\eqref{eq:nwlayers-bound}. This yields 195 $w$-projection planes or $w$-layers at 10\degree zenith angle. Some configurations with high $w$-values fail to image with CASA, presumably because the $w$-kernels become to large.

TODO discuss CASA version

TODO discuss desktop computer specs
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{llll}%
\hline\hline
& \textbf{WSClean} & \textbf{WSClean}    & \textbf{CASA} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{1}{l}{\textit{Zenith angle 0\degree}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & & 1.90 mJy \\
Computational time & 8 min & & 19.3 min \\
\hline
\multicolumn{1}{l}{\textit{Zenith angle 10\degree}} \\
\hline
Source flux standard error & 1.75\% & & 2.41 \%\\
RMS in residual image & 0.88 mJy & & 1.04 mJy \\
Computational time & 15.3 min & & 178.2 min \\
\hline\hline
\end{tabular}\end{center}\end{table}

\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/residual-casa}
\includegraphics[width=8cm]{img/residual-wsclean}
\caption{Residual images after Cotton-Schwab cleaning of a simulated MWA observation, using similar inversion and cleaning parameters. Left image: w-projection using CASA; right image: w-stacking using WSClean. The image has been zoomed in just above the image centre, with the image centre being displayed in the bottom-centre of the panels. Sources in the image produced with CASA are less accurately subtracted, leading to a residual noise level of 1.9 mJy, whereas WSClean reaches 0.94 mJy. Since the effect is stronger away from the phase centre, we believe this is caused by the finite size of the w-kernels used in w-projection, which leads to inaccuracies.}
\label{fig:residuals}
\end{center}
\end{figure*}

We performed a few measurements to validate that there are no major accuracy difference between WSClean and CASA. To do so, we have simulated a MWA observation with hundred sources of 1~Jy in a 20 degrees area. A unitary primary beam is assumed. We image the set both with WSClean and CASA, using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. Aegean (TODO discuss) was subsequently used to perform source detection on the produced images. Even though no noise is added to the visibilities, sidelobe noise of the residual 10~mJy noises triggers a few false detections, which are ignored.

Table~\ref{tbl:accuracy-measurements} lists the measured standard error of the source brightnesses as detected by Aegean, as well as the RMS in the residual image.

TODO show aliasing errors for too few wlayers.

\subsection{Performance}
We measured the performance of the imagers using several MWA data sets. Each specific configuration is run 5 times and standard deviations are calculated over these 5 runs. The variations between runs is typically a few seconds. In each benchmark, the wall-clock time is measured that is required to produce the synthesized point-spread function and the image itself. No cleaning or prediction was performed.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-zenith-angle/za}
\caption{Effect of zenith angle (ZA) on imaging performance. Image width and height are $2048$ for the dashed line and $3072$ for the solid line. For the latter, CASA fails to image observations with ZA $\ge 15$. ZA dramatically affects the computational cost of w-projection, and WSClean is more than an order of magnitude faster at ZA $\ge$ 10\degree with a $3072$ image size (or ZA $\ge$ 30\degree at $1536$.) }
\label{fig:timing-zenith-angle}
\end{center}
\end{figure}

@@

\subsection{Cleaning a w-stacked image}
TODO describe how to do
TODO \citet{hogbom-clean}

It is tempting to use a very low number of $w$-layers, and define the major clean iteration threshold such that it will never clean any $w$-aliasing artefacts. However, predicting the visibilities from an image with a similar low number of $w$-layers will cause similar aliased errors, and thus will subtract incorrect fluxes from the visibilities. In the end, even though the sources causing the $w$-aliasing effects are removed, aliasing artefacts will still remain.

@@

\section{Discussion} \label{sec:discussion}
TODO: describe hybrid between w-stacking, w-projection and w-snapshots (if not yet discussed in Cornwell's papers)

>>>>

\subsection{Discretization of w} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers. WSClean was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle. This would normally require imaging with $\sim195$ $w$-layers. Original sources are 1~Jy, ghost sources are approximately 20\%. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce ghost sources by themselves. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal TODO in contrast to some article? -- could be explained by having more samples at smaller $w$s. From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. [TODO table of arrays with $w$ values at zenith/more angles/more freq?] For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that, in the case of the MWA, tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

Using too few $w$-layers results in decorrelation of off-axis sources in the longer baselines. Additionally, $w$-aliasing errors can cause ghost sources in the image. An (extreme) example of this effect is shown in Fig.~\ref{fig:aliasing-example}. When Cotton-Schwab cleaning includes prediction with too few $w$-layers, incorrect values will be subtracted from the visibilities even when no aliased sources are cleaned during minor iterations. Therefore, accurate prediction is more important than accurate imaging, because aliasing artefacts are attenuated by cleaning as long as the model is subtracted accurately. Possible ways to allow less $w$-layers during imaging is by predicting with more $w$-layers or a direct Fourier transform. We have not looked into these options yet.

@@

\begin{table}
 \caption{Scaling of the computational cost for various imaging steps} \label{tbl:computational-cost-per-operation}
 \begin{tabular}{lll}
   \textbf{Operation} & \textbf{W-stacking} & \textbf{W-projection} \\
   \hline\hline
   Fourier transform(s) & $N_\textrm{wlay} N^2_\textrm{pix} \log N_\textrm{pix}$ & $N^2_\textrm{pix} \log N_\textrm{pix}$ \\
   W-term corrections   & $N_\textrm{wlay} N^2_\textrm{pix}$ & $N_\textrm{vis} N^2_\textrm{wkern}$ \\
   Gridding & $N_\textrm{vis}N^2_\textrm{kern}$ & $N_\textrm{vis} N_\textrm{kern}^2$ \\
   \hline\hline
 \end{tabular}
\end{table}

@@

The new imager, called ``WSClean'', is written in the C++ language. It reads visibilities from CASA measurement sets and writes output images to Flexible Image Transport System (FITS) files. Several steps are multi-threaded: reading and writing; gridding from and to different $w$-layers; performing the Fast Fourier transforms; and H\"ogbom Clean iterations (peak-finding and image subtraction, \citealt{hogbom-clean}). For the latter, intrinsics are used as well. Because of these optimizations, minor Cleaning iterations are performed at a rate of hundreds per second at $3072\times3072$. This is fast enough to make the Clark Clean optimization \citep{clark-clean}, which consists of considering only a subset of pixels with a reduced PSF, less relevant. However, the PSF varies for imaging with non-zero $w$-values, and subtracting a constant PSF in image space leads to inaccuracies. After a number of minor iterations, it is therefore beneficial to invert the model back to the visibilities via prediction, and subtract the model directly from the visibilities \citep{wprojection-cornwell}. This is similar to the Cotton-Schwab cleaning method \citep{cotton-schwab-clean}.

@@

\subsection{Full-sky imaging}
For low-frequency telescopes, it can be of interest to do full-sky imaging. In the case of MWA, the tile beam can have strong sidelobes at a distance of more than 90\degree from the pointing centre. Imaging these sidelobes might be relevant because it might either be scientifically interesting (e.g. when searching for transients), or be useful for self-calibration or full-sky deconvolution. For example, the latter is useful when a strong source in a sidelobe affects the sensitivity near the pointing centre.

To efficiently image the full sky, an observation is split into short snapshots, their phase centre is changed to zenith and subsequently a suitable area is imaged. Depending on desired resolution and dynamic range, this might require very large images. To make an image at the MWA resolution, the image needs to have approximately 10K pixels in width and height. The $w$-values will be very small at zenith, because at zenith only the vertical offsets of the antennas will contribute to the $w$-value. For the MWA, the maximum height difference between tiles is $8.5$m. The MWA tiles are in fact on a slight slope, and by fitting a plane to the antennas and changing the phase centre towards the normal of that plane, the maximum $w$-value decreases to $5.5\textrm{m} / \lambda$. In the following sections, when discussing the MWA zenith direction we are referring to this optimal $w$-direction.

\subsection{W-snapshots} \label{sec:snapshot-imaging}
\begin{figure*}
\begin{center}
\includegraphics[width=8cm]{img/vela-normal-projection}
\includegraphics[width=8cm]{img/vela-zenith-projection}
\caption{13-min MWA observation of supernova remnants Vela and Puppis A. Left: normal projection centred on Puppis A, right: phase-rotated to zenith to reduce $w$-values, recentred on Puppis A during imaging. Both Stokes-I images have been made with WSClean using the Cotton-Schwab clean algorithm. No beam correction was applied. TODO time difference and more discussion, replace with deeper images.}
\label{fig:vela-projection-example}
\end{center}
\end{figure*}
As shown, all-sky imaging can be quite efficient with snapshot imaging towards zenith. However, if one is not interested in imaging the whole sky, and the direction of interest is far from zenith, the computational overhead of making all-sky images is undesirable.

Instead of imaging the whole sky, the image can be recentred from $(l,m)$ to $(\hat{l},\hat{m})=(l+\Delta l,m+\Delta m)$ by performing the substitution $(l,m)\leftarrow(\hat{l},\hat{m})$ in Eq.~\eqref{eq:wstacking}:
%When the Fourier transform is performed, the $w$ term has been corrected (either by w-stacking or by w-projection). Therefore, if $U(u, v)$ is the input to the 2-dimensional Fourier transform, the transform is described by
\begin{align}\notag
\frac{I'(\hat{l}, \hat{m})\left(w_e - w_0\right)}{\sqrt{1-\hat{l}^2-\hat{m}^2}} = \int\limits_{w_0}^{w_e} e^{2\pi i w(\sqrt{1-\hat{l}^2-\hat{m}^2}-1)} \cdot \\
\iint e^{2 \pi i \left(u\Delta l + v\Delta m\right)} V(u,v,w) e^{2\pi i \left(ul + vm\right)} du dv dw.
\end{align}
In words, recentering an image involves accounting for the position shift during the $w$-correction and final scaling, and shifting the visibilities in phase by multiplication with $e^{2 \pi i \left(u\Delta l + v\Delta m\right)}$ prior to the imaging. This essentially is a method of implementing the $w$-snapshots algorithm as described by \citet{widefield-imaging-ska-cornwell}. By doing the inverse corrections during prediction, such a shifted visibility set can be cleaned with Cotton-Schwab iterations similar to a non-shifted set.

@@

Another method to shift an image is by using the periodicity of the FFT function. Since sources outside the field of view will be aliased back into the field, the image can be transformed to have the alias in the centre. This complicates gridding during imaging, because the antialiasing kernel needs to have a pass-band shape instead of a low-pass shape. Therefore, we chose to implement the former method.

\section{Analysis} \label{sec:analysis}
We will now analyse the performance and accuracy of our implementation, and compare it with the w-projection implementation in CASA. For the analysis, we use imaging parameters which are common for MWA Imaging. When a specific parameter value is not mentioned, the settings from Table~\ref{tbl:default-parameters} are used. The number of $w$-projection planes in w-projection is kept equal to the number of $w$-layers in w-stacking, and is set to the right hand side of Eq.~\eqref{eq:nwlayers-bound}. This yields 195 $w$-planes/layers at 10\degree zenith angle. Some configurations with high $w$-values fail to image with CASA, presumably because the $w$-kernels become to large.

In all analyses CASA stable release 42.0, revision 26465 was used, which was released September 2013. It includes performance improvements to the \texttt{clean} task which have recently been made. The latest version as of February 2014 of WSClean was used. The tests were run on a high-end desktop, with a 3.20-GHz Intel Core i7-3930K processor and 32 GB of memory. The data are stored on a multi-disk array with 5 spinning hard disks, which has a combined read rate of about 450 MB/s. 
\begin{table}%
\caption{Parameter values used during benchmarks, unless otherwise mentioned.} \label{tbl:default-parameters}%
\begin{center}\begin{tabular}{rl}%
\hline
Array & MWA 128T \\
Image size & $3072 \times 3072$ \\
Angular resolution & $0.72'$ \\
Number of visibilities & $3.5 \times 10^8$ \\
Time resolution & 2~s \\
Frequency resolution & 40~kHz \\
Observation duration & 112 s\\
Bandwidth & 30.72 MHz (768 channels)\\
Central frequency & 182 MHz \\
Zenith angle at phase centre & 10\degree \\
Max $w$-value for phase centre & 172 $\lambda$ (283 m) \\
Nr. polarizations in set & 4 \\
Imaged polarization & XX \\
Imaging mode & Multi-frequency synthesis \\
Weighting & uniform \\
Data size & 18 GB \\
\hline
\end{tabular}\end{center}\end{table}

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{lrrr}%
\hline\hline
& \textbf{WSClean} & \textbf{WSClean}    & \textbf{CASA} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (minimum w-layers/planes)}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & --- & 1.90 mJy\\
Computational time & 8.5 min & & 19.3 min\\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (128 w-layers/planes)}} \\
\hline
Source flux standard error & 1.39\% & & 2.08\% \\
RMS in residual image & 0.94 mJy & --- & 0.94 mJy \\
Computational time & 10.3 min & & 19.6 min \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 10\degree}} \\
\hline
Source flux standard error & 1.75\% & 2.92\% & 2.41 \%\\
RMS in residual image & 0.90 mJy & 1.05 mJy & 1.07 mJy \\
Computational time & 15.3 min & 6.6 min & 178.2 min \\
& & recentreTODO & \\
\hline\hline
\end{tabular}\end{center}\end{table}

\begin{figure*}
\begin{center}
\includegraphics[height=5cm]{img/residual-casa}\hspace{1cm}\includegraphics[height=5cm]{img/residual-wsclean}
\caption{Residual images after Cotton-Schwab cleaning of a simulated 10\degree-zenith angle MWA observation using CASA (left) and WSClean (right), using similar inversion and cleaning parameters. The panels show a small part of the full images. The full field contains 100 simulated sources over 20\degree. Sources in the image produced with CASA are slightly less accurately subtracted, leading to a residual noise level of 1.07 mJy and some visible artefacts, whereas WSClean reaches 0.90 mJy. Since the effect is stronger away from the phase centre, it is likely that this is caused by the finite size of the w-kernels used in w-projection, which leads to inaccuracies.}
\label{fig:residuals}
\end{center}
\end{figure*}

We performed measurements to assess the accuracy of WSClean and CASA. To do so, we have simulated a MWA observation without system noise, but with hundred sources of 1~Jy in a 20\degree diameter area. A unitary primary beam is assumed. We imaged the simulated set both with WSClean and CASA, using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. The Aegean program \citep{aegean-hancock-2012} was used to perform source detection on the produced images. Sidelobe noise of the residual 10~mJy noise triggers a few false detections. These are ignored.

Table~\ref{tbl:accuracy-measurements} lists the measured standard errors of the source brightnesses as detected by Aegean and the RMS in the residual image. In all tests, WSClean performs slightly more accurate. At zenith, CASA produces a residual RMS twice as high as WSClean. This is caused by the fact that we keep the number of w-projection planes in w-projection equal to the number of w-layers in w-stacking, resulting in only 12 $w$-planes at zenith. This evidently has a stronger effect on the w-projection algorithm. However, the computational performance of the w-projection algorithm is hardly affected by the number of w-projection planes, and in practical situations one would always use more w-projection planes\footnote{In the zenith case CASA actually issues a warning to use more $w$-projection planes.}. When 128 w-projection planes are used in CASA, its residual RMS is equal to WSClean with 12 $w$-layers. We do not know why this parameter needs to be higher in CASA to reach the same RMS. In theory it should be enough, since Eq.~\eqref{eq:nwlayers-bound} results in $N_\textrm{wlayers}\gg 6$ for the top-left image pixel that is 26\degree away from the centre, while the sources are all within 10\degree. Also unexpected is that the source flux density becomes worse by increasing the number of planes. For WSClean both values stay approximately the same when the number of $w$-layers is increased.

In the $ZA=10\degree$ case, CASA is slightly less accurate. This is consistent, and can actually be observed in the images as shown in Fig.~\ref{fig:residuals}. TODO this might be caused by a bug; check: We have also made an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is lower compared to the normal projection. This is due to the fact that the sources are further away from the phase centre, and sources have therefore on average higher $(l,m)$-values, thereby increasing the average $w$-correction. Nevertheless, the reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

WSClean is faster in all tested cases. WSClean and CASA determine a different next major iteration threshold. CASA performs 5 major iterations at zenith and 8 at 10\degree angle, while WSClean always performs 5 major iterations. Major iterations are dominating the computing time. We will look more closely at the difference in performance in the next section.

Many parameters affect the accuracy more than the observed differences between the methods. These include the number of major iterations, $w$-planes/layers, restoring beam, image size, weighting, kernel size and grid oversampling. Therefore, the differences are not very significant. The results do show that synthesis imaging using common imaging parameters results in an approximate flux-density standard error of at least $\sim1\%$.

\subsection{Performance analysis}
We measured the performance of the imagers using several MWA data sets. Each specific configuration is run 5 times and standard deviations are calculated over these 5 runs. The variations between runs is typically a few seconds. In each benchmark, the wall-clock time is measured that is required to produce the synthesized point-spread function and the image itself. No cleaning or prediction was performed.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/benchmark-zenith-angle/za}
\caption{Effect of zenith angle (ZA) on imaging performance. Image width and height are $2048$ for the dashed line and $3072$ for the solid line. CASA fails to image observations with ZA $\ge 15$ with a $3072$ image size. ZA dramatically affects the computational cost of w-projection. }
\label{fig:timing-zenith-angle}
\end{center}
\end{figure}

@@

@@

\section{Discussion} \label{sec:discussion}
WSClean has been build out of necessacity to be able to make images with the MWA.

TODO: describe hybrid between w-stacking, w-projection and w-snapshots (if not yet discussed in Cornwell's papers)
